---
title: "model stability plots, cross validation"
author: "SID: 490300286"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---

```{r setup, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(factoextra)
library(knitr)
library(kableExtra)
library(gridExtra)
library(broom)
library(fpc)
library(sjPlot)
library(caret)
library(ggfortify)
library(ggforce)
library(Rtsne)
library(mplot)
set.seed(2000)
load('tech_data_biom_and_nutr.Rdata')
```



# creating final dataset

```{r}
# selecting macros
foods = tech_nutr %>% dplyr::select(ABSPID,
                             VEGLEG1N, 
                             VEGLEG2N,
                              FRUIT1N,
                              FRUIT2N,
                              DAIRY1N,
                              DAIRY2N,
                              MEAT1N,
                              MEAT2N,
                              GRAINS1N,
                              GRAINS2N)


# gettin average of macros
avg_veges <- rowMeans(foods[ , c(2,3)], na.rm=TRUE)
avg_fruit <- rowMeans(foods[ , c(4,5)], na.rm=TRUE)
avg_dairy <- rowMeans(foods[ , c(6,7)], na.rm=TRUE)
avg_meat <- rowMeans(foods[ , c(8,9)], na.rm=TRUE)
avg_grains <- rowMeans(foods[ , c(10,11)], na.rm=TRUE)


dat <- cbind(tech_nutr$ABSPID, 
             avg_veges,
             avg_fruit,
             avg_dairy,
             avg_meat,
             avg_grains)

dat <- as_tibble(dat)
colnames(dat)[1] <- "ABSPID"

dat$avg_veges <- as.numeric(dat$avg_veges)
dat$avg_fruit <- as.numeric(dat$avg_fruit)
dat$avg_dairy <- as.numeric(dat$avg_dairy)
dat$avg_meat <- as.numeric(dat$avg_meat)
dat$avg_grains <- as.numeric(dat$avg_grains)



tech_biom1 = tech_biom %>% dplyr::select(c(1:53))

final = merge(dat, tech_biom1, by = "ABSPID")
final = final %>% dplyr::select(avg_veges,
                         avg_fruit,
                         avg_dairy,
                         avg_meat,
                         avg_grains,
                         BMISC,
                         AGEC,
                         SEX,
                         PHDKGWBC,
                         PHDCMHBC,
                         PHDCMWBC,
                         SF2SA1QN)


final$w2hratio = (final$PHDCMWBC)/(final$PHDCMHBC)

# to make results more robust, focus on adults only

final <- final %>% filter(AGEC > 17)

final = final %>% na.omit()
```

The code in this code chunk just makes a final data frame with the avg servings of each food group for each participant

# short analysis


## linear regression with BMI as outcome

```{r}
bmi_full <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
bmi_null <- lm(BMISC ~ 1, dat = final)

```


## linear regression with waist to heigh ratio as outcome

```{r}
w2hratio_full <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
w2hratio_null <- lm(w2hratio ~ 1, dat = final)
```

## linear regression with waist circumference as outcome

```{r}
waist_full <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final) 
waist_null <- lm(PHDCMWBC ~ 1, dat = final) 

```


NOTE: robust regression was attempted and gave similar results to OLS, so just using OLS instead.


## table of all models

```{r full_models, cache = T}
full_model_table <- tab_model(bmi_full, waist_full, w2hratio_full, show.ci = F, show.se = T)
full_model_table
```


dairy is insignificant for all


## Model selection and performance (10 fold CV)

Model selection using backward and forward selection

### BMI

```{r}
n = nrow(final)

# bmi_AIC is just full model
# bmi_AIC <- step(bmi_full, direction = "backward", trace = F)

bmi_BIC <- step(bmi_full, direction = "backward", trace = F, k = log(n))

bmi_BIC %>% summary()

# fwd AIC model same as back
# step(bmi_null, scope = list(lower = bmi_null, upper = bmi_full), direction = "forward", trace = F)

```

### waist circumeference

```{r}
waist_AIC <- step(waist_full, direction = "backward", trace = F)

# waist_BIC same as AIC
# waist_BIC <- step(waist_full, direction = "backward", trace = F, k = log(n))

waist_AIC %>% summary()
# backward selection same as fwd
# step(waist_null, scope = list(lower = waist_null, upper = waist_full), direction = "forward", trace = F)
```

### w2h ratio


```{r}
w2hratio_AIC <- step(w2hratio_full, direction = "backward", trace = F)
w2hratio_BIC <- step(w2hratio_full, direction = "backward", trace = F, k = log(n))

w2hratio_AIC %>% summary()
w2hratio_BIC %>% summary()
# backward selection same as fwd
# step(w2hratio_null, scope = list(lower = w2hratio_null, upper = waist_full), direction = "forward", trace = F)
```

# model stability plots

```{r mplots, cache = T}
vis.obj_bmi <- vis(bmi_full, B = 150, redundant = TRUE, nbest = "all",
               seed = 2021)
vis.obj_waist <- vis(waist_full, B = 150, redundant = TRUE, nbest = "all",
               seed = 2021)
vis.obj_w2hratio <- vis(w2hratio_full, B = 150, redundant = TRUE, nbest = "all",
               seed = 2021)


```


## BMI

```{r}
plot(vis.obj_bmi, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "boot")

plot(vis.obj_bmi, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "vip")

```


Shows that dairy is close to the Redundant variable (RV) curve, hence dairy does not add any significant information to the model.

# w2h ratio

```{r}
plot(vis.obj_w2hratio, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "boot")

plot(vis.obj_w2hratio, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "vip")
```

Shows that dairy is close to the Redundant variable (RV) curve, hence dairy does not add any significant information to the model.


# waist circumference


```{r}
plot(vis.obj_waist, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "boot")

plot(vis.obj_waist, 
     interactive = FALSE, 
     highlight = "avg_dairy", 
     which = "vip")
```

Shows that dairy is close to the Redundant variable (RV) curve, hence dairy does not add any significant information to the model.



From all of these above plots, we can see that models containing `avg_dairy` are only dominant when there are a large number of parameters in the model (i.e. close to the full model). From this, we can see that avg_dairy is not particularly stable, and therefore suggests it should not be included in a final model.


# ANOVA of models

## BMI

```{r}
anova(bmi_BIC, bmi_full)
```

From this, we can see that when we can see that p = 0.141 > 0.05 when dairy is added as a covariate. Hence it should not be included in the final model.

## Waist circumference

```{r}
anova(waist_AIC, waist_full)
```

From the above ANOVA, p > 0.05, hence model with dairy is insignificant so final model should exlcude dairy


```{r}


anova(w2hratio_BIC, w2hratio_AIC, w2hratio_full)

```

```{r}
anova(w2hratio_BIC, w2hratio_AIC, w2hratio_full)

```

The BIC is nested within the AIC model which is nested within the full model. From the output, the ANOVA table shows p < 0.05 for the AIC model only. Hence all food groups except dairy are significant. 


# Cross validation


```{r}
params = trainControl(method = "cv", number = 10, verboseIter = FALSE)

set.seed(2021)

cv_objects = list(
  bmi_full = train(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat +
    avg_grains + AGEC + SEX, 
    method = "lm", 
    data = final,
    trControl = params),
  bmi_BIC = train(BMISC ~ avg_veges + avg_fruit + avg_meat + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),
    bmi_tree = train(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
                   data = final,
                   method = 'rpart',
                   tuneLength  = 15, 
                   trControl = params),
  waist_full = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),  
  waist_AIC = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params), 
  waist_tree = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
                   data = final,
                   method = 'rpart',
                   tuneLength  = 15, 
                   trControl = params),
  w2hratio_full = train(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_AIC = train(w2hratio ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_BIC = train(w2hratio ~ avg_veges + avg_fruit + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),
  w2hratio_tree = train(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
                   data = final,
                   method = 'rpart',
                   tuneLength  = 15, 
                   trControl = params)
)

cv_results_bmi = resamples(cv_objects[1:3], metric = "RMSE")
ggplot(cv_results_bmi) +
  theme_bw() +
  labs(x = "Models", y = "RMSE", title = "10-Fold CV Performance")

cv_results_waist = resamples(cv_objects[4:6], metric = "RMSE")
ggplot(cv_results_waist) +
  theme_bw() +
  labs(x = "Models", y = "RMSE", title = "10-Fold CV Performance")


cv_results_w2hratio = resamples(cv_objects[7:10], metric = "RMSE")
ggplot(cv_results_w2hratio) +
  theme_bw() +
  labs(x = "Models", y = "RMSE", title = "10-Fold CV Performance")



```
Looking at RMSE of BMI models, all have very similar 



## comaring standardized betas

```{r}
# creating scaled df
sex <- final$SEX
socioeconomic <- final$SF2SA1QN
scaled_df <- final %>% 
  dplyr::select(-c(SEX, SF2SA1QN)) %>% 
  scale() %>% as.data.frame()


scaled_df$SEX <- sex
scaled_df$SF2SA1QN <- socioeconomic

# lm scaled to get standardized betas to compare
bmi_scaled <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
w2h_scaled <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
waist_scaled <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)

```

```{r st_betas, cache = T}
st_beta_table <- tab_model(bmi_scaled, w2h_scaled, waist_scaled)
st_beta_table
```
meat has the biggest effect on BMI overall effect and increases BMI.
fruit and veges decrease waist and waist to height ratio.

