---
title: "Plan"
author: "SID: 490300286"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---


# summary of plan

we investigate the effect of diet composition on different measures of obesity.


Obesity can be measured in multiple ways

* BMI
* Waist circumference (https://www.nature.com/articles/s41574-019-0310-7)
* Waist to height ratio (https://bmcmedicine.biomedcentral.com/articles/10.1186/s12916-014-0207-1)

But each of these may categorise obesity differently.

In terms of diet composition, we look at 

* daily servings of meat
* daily servings of veg
* daily servings of dairy
* daily servings of fruit
* daily servings of grains

we then want to see what is the effect of each food type on a measurement for obesity (whether it is significant or not), and compare how this varies across each obesity measurement.

> e.g. Maybe meat servings is a significant contributor to BMI but not to waist circumference.

Then, we can compare to see the effect of each food type across different obesity measures. Examples

> e.g. Does fruit servings have a bigger effect on BMI, waist to height ratio, or waist circumference
> which of the food types has the biggest contribution to BMI reduction/gain.

We can do this by calculating and comparing the standardized betas.

(we can adjust for obvious contributors to each measurement such as age and sex). 


# exec summary of findings

- dairy is insignificant conributor to all measures of obesity


```{r setup, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(factoextra)
library(knitr)
library(kableExtra)
library(gridExtra)
library(broom)
library(fpc)
library(sjPlot)
library(caret)
library(ggfortify)
library(ggforce)
library(Rtsne)
library(e1071)
set.seed(2000)
load('tech_data_biom_and_nutr.Rdata')
```


```{r}
# selecting macros
foods = tech_nutr %>% dplyr::select(ABSPID,
                             VEGLEG1N, 
                             VEGLEG2N,
                              FRUIT1N,
                              FRUIT2N,
                              DAIRY1N,
                              DAIRY2N,
                              MEAT1N,
                              MEAT2N,
                              GRAINS1N,
                              GRAINS2N)

# gettin average of macros
avg_veges <- rowMeans(foods[ , c(2,3)], na.rm=TRUE)
avg_fruit <- rowMeans(foods[ , c(4,5)], na.rm=TRUE)
avg_dairy <- rowMeans(foods[ , c(6,7)], na.rm=TRUE)
avg_meat <- rowMeans(foods[ , c(8,9)], na.rm=TRUE)
avg_grains <- rowMeans(foods[ , c(10,11)], na.rm=TRUE)


dat <- cbind(tech_nutr$ABSPID, 
             avg_veges,
             avg_fruit,
             avg_dairy,
             avg_meat,
             avg_grains)

dat <- as_tibble(dat)
colnames(dat)[1] <- "ABSPID"

dat$avg_veges <- as.numeric(dat$avg_veges)
dat$avg_fruit <- as.numeric(dat$avg_fruit)
dat$avg_dairy <- as.numeric(dat$avg_dairy)
dat$avg_meat <- as.numeric(dat$avg_meat)
dat$avg_grains <- as.numeric(dat$avg_grains)



tech_biom1 = tech_biom %>% dplyr::select(c(1:53))

final = merge(dat, tech_biom1, by = "ABSPID")
final = final %>% dplyr::select(avg_veges,
                         avg_fruit,
                         avg_dairy,
                         avg_meat,
                         avg_grains,
                         BMISC,
                         AGEC,
                         SEX,
                         PHDKGWBC,
                         PHDCMHBC,
                         PHDCMWBC,
                         SF2SA1QN)


final$w2hratio = (final$PHDCMWBC)/(final$PHDCMHBC)

# to make results more robust, focus on adults only

final <- final %>% filter(AGEC > 17)

final = final %>% na.omit()
```


## IDA on varibales in df

```{r, message = F, warning = F}
numeric_hist <- function(data, x) {
  ggplot(data, aes_string(x = `x`)) +
  geom_histogram()
}


numeric_hist(final, x = "BMISC")
numeric_hist(final, x = "avg_dairy")
numeric_hist(final, x = "avg_fruit")
numeric_hist(final, x = "avg_meat")
numeric_hist(final, x = "avg_grains")
numeric_hist(final, x = "avg_veges")
numeric_hist(final, x = "AGEC")
```


# short analysis


## linear regression with BMI as outcome

```{r}
bmi_full <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
bmi_null <- lm(BMISC ~ 1, dat = final)

```


## linear regression with waist to heigh ratio as outcome

```{r}
w2hratio_full <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
w2hratio_null <- lm(w2hratio ~ 1, dat = final)
```

## linear regression with waist circumference as outcome

```{r}
waist_full <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final) 
waist_null <- lm(PHDCMWBC ~ 1, dat = final) 

```


NOTE: robust regression was attempted and gave similar results to OLS, so just using OLS instead.


## table of all models

```{r}
tab_model(bmi_full, waist_full, w2hratio_full, show.ci = F, show.se = T,
            dv.labels = c("BMI", "Waist circumf.", "waist to height ratio"))
```


dairy is insignificant for all


## Model selection and performance (10 fold CV)

Model selection using backward and forward selection

### BMI

```{r}
n = 9979

# bmi_AIC is just full model
# bmi_AIC <- step(bmi_full, direction = "backward", trace = F)

bmi_BIC <- step(bmi_full, direction = "backward", trace = F, k = log(n))

# fwd AIC model same as back
# step(bmi_null, scope = list(lower = bmi_null, upper = bmi_full), direction = "forward", trace = F)

```

### waist circumeference

```{r}
waist_AIC <- step(waist_full, direction = "backward", trace = F)

# waist_BIC same as AIC
# waist_BIC <- step(waist_full, direction = "backward", trace = F, k = log(n))


# backward selection same as fwd
# step(waist_null, scope = list(lower = waist_null, upper = waist_full), direction = "forward", trace = F)
```

### w2h ratio


```{r}
w2hratio_AIC <- step(w2hratio_full, direction = "backward", trace = F)
w2hratio_BIC <- step(w2hratio_full, direction = "backward", trace = F, k = log(n))

# backward selection same as fwd
# step(w2hratio_null, scope = list(lower = w2hratio_null, upper = waist_full), direction = "forward", trace = F)
```




```{r}
params = trainControl(method = "cv", number = 10, verboseIter = FALSE)

set.seed(2021)

cv_objects = list(
  bmi_full = train(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat +
    avg_grains + AGEC + SEX, 
    method = "lm", 
    data = final,
    trControl = params),
  bmi_BIC = train(BMISC ~ avg_veges + avg_fruit + avg_meat + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),
  waist_full = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),  
  waist_AIC = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params), 
  w2hratio_full = train(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_AIC = train(w2hratio ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_BIC = train(w2hratio ~ avg_veges + avg_fruit + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params)   
)

cv_results_bmi = resamples(cv_objects[1:2], metric = "Rsquared")
ggplot(cv_results_bmi) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")

cv_results_waist = resamples(cv_objects[3:4], metric = "Rsquared")
ggplot(cv_results_waist) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")


cv_results_w2hratio = resamples(cv_objects[5:7], metric = "Rsquared")
ggplot(cv_results_w2hratio) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")
```

## Regression assumptions

#### First model (full BMI)

```{r}
bmi_full <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat +
  avg_grains + AGEC + SEX, 
  method = "lm", 
  data = final,
  trControl = params)
summary(bmi_full)
model.diag.metrics <- augment(bmi_full)
head(model.diag.metrics)
```

```{r}
plot1 <- ggplot(model.diag.metrics, aes(avg_meat, BMISC)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = avg_meat, yend = .fitted), color = "red", size = 0.3)
plot2 <- ggplot(model.diag.metrics, aes(avg_dairy, BMISC)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = avg_dairy, yend = .fitted), color = "red", size = 0.3)
plot3 <- ggplot(model.diag.metrics, aes(avg_grains, BMISC)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = avg_grains, yend = .fitted), color = "red", size = 0.3)
plot4 <- ggplot(model.diag.metrics, aes(avg_veges, BMISC)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = avg_veges, yend = .fitted), color = "red", size = 0.3)


grid.arrange(plot1, plot2, plot3, plot4, ncol=2)
```

The residuals mostly seem to follow the same pattern across features


```{r}
par(mfrow=c(2,2))
plot(bmi_full)
```

Residuals vs. fitted is relatively linear, indicating a linear relationship. 
Normal Q-Q plot, standardised residuals follows closely the normal line, indicating normality of residuals.
Scale-location has no obvious non-linear pattern, indicating homoscedasticity.
In the residuals vs. leverage graph a number of points are outside of Cook's distance, indicating they are outliers. Some outlier effects may exist in the regression models specified above.



## Comparing standardized betas

```{r}
# creating scaled df
sex <- final$SEX
socioeconomic <- final$SF2SA1QN
scaled_df <- final %>% 
  dplyr::select(-c(SEX, SF2SA1QN)) %>% 
  scale() %>% as.data.frame() 


scaled_df$SEX <- sex
scaled_df$SF2SA1QN <- socioeconomic

# lm scaled to get standardized betas to compare
bmi_scaled <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
w2h_scaled <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
waist_scaled <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)


tab_model(bmi_scaled, w2h_scaled, waist_scaled)
```
meat has the biggest effect on BMI overall effect and associated with increased BMI.
fruit and veges associated with reduced waist circumference and waist to height ratio.

## PCA and clustering to see any trends in obesity


### Performing PCA
```{r}

food_groups <- final %>% dplyr::select(avg_dairy,
                                avg_fruit,
                                avg_grains,
                                avg_meat,
                                avg_veges)

numerics <- final %>% dplyr::select(-SEX,
                                    -SF2SA1QN)

# principal component analysis
res.pca <- prcomp(
  food_groups,
  center = T,
  scale = T
)

# extracting pc1 and pc2
pcs = res.pca$x[, c(1,2)] %>% as.data.frame()

```

### Clustering

```{r}
## keeping original data
pcs1 <- pcs
## Creating k-means clustering model
fit_cluster_kmeans_pca <- kmeans(scale(pcs), 5)
# Assigning the result to the data used to create the tsne
pcs1$cl_kmeans <- factor(fit_cluster_kmeans_pca$cluster)
# Creating hierarchical cluster model
fit_cluster_hierarchical_pca <- hclust(dist(scale(pcs)), method = "ward.D2")
#  Assigning the result to the data used 
pcs1$cl_hierarchical <- factor(cutree(fit_cluster_hierarchical_pca, k=5))
```

### Hierarchical clustering visualisation
```{r}
# obesity measured via BMI >= 30
pcs1$BMISC <- final$BMISC
pcs1$obese_BMI <- pcs1$BMISC >= 30

# obesity measured via waist circumference

final$obese_waist_circum = NA

for (i in 1:nrow(final)) {
  if (final$SEX[i] == "2") {
    if (final$PHDCMWBC[i] > 88) {
      final$obese_waist_circum[i] = 1
    } else {
      final$obese_waist_circum[i] = 0
    }
    
  } else if (final$SEX[i] == "1") {
    if (final$PHDCMWBC[i] > 102) {
      final$obese_waist_circum[i] = 1
    } else {
      final$obese_waist_circum[i] = 0
    }
    
  }

}

pcs1$obese_waist_circum <- final$obese_waist_circum


# obesity measured using w2hratio


final$obese_w2hratio = NA

for (i in 1:nrow(final)) {
  if (final$SEX[i] == "2") {
    if (final$w2hratio[i] >= 0.58) {
      final$obese_w2hratio[i] = 1
    } else {
      final$obese_w2hratio[i] = 0
    }
    
  } else if (final$SEX[i] == "1") {
    if (final$w2hratio[i] >= 0.63) {
      final$obese_w2hratio[i] = 1
    } else {
      final$obese_w2hratio[i] = 0
    }
    
  }

}

pcs1$obese_w2hratio <- final$obese_w2hratio
```

### Determine the ideal number of clusters
#### k-means
```{r}
data <- scale(pcs)
k.max <- 15
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=50,iter.max = 15 )$tot.withinss})
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares", main="k-means elbow")
```
#### hierarchical
```{r}
#get dendrogram
plot(fit_cluster_hierarchical_pca, hang = -1)
```



```{r}
pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=obese_BMI),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_hierarchical,
                        label=cl_hierarchical),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("Hierarchical, BMI")


pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=factor(as.character(obese_waist_circum))),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_hierarchical,
                        label=cl_hierarchical),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("Hierarchical, waist circumference")


pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=obese_w2hratio %>% as.character() %>% as.factor()),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_hierarchical,
                        label=cl_hierarchical),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("Hierarchical, waist to height ratio")
```


```{r}
pcs1 %>% group_by(cl_hierarchical) %>% 
  summarise(prop_obese_BMI = sum(as.numeric(obese_BMI))/n())

pcs1 %>% group_by(cl_hierarchical) %>% 
  summarise(prop_obese_waist_circum = sum(as.numeric(obese_waist_circum))/n())


pcs1 %>% group_by(cl_hierarchical) %>% 
  summarise(prop_obese_w2hratio = sum(as.numeric(obese_w2hratio))/n())
```



```{r}
scaled_numerics <- scale(numerics)
scaled_numerics <- as.data.frame(scaled_numerics)
scaled_numerics$cl_hierarchical <- pcs1$cl_hierarchical


scaled_numerics %>% group_by(cl_hierarchical) %>% 
  summarise(avg_veges = mean(avg_veges) %>% round(2),
            avg_dairy = mean(avg_dairy) %>% round(2),
            avg_fruit = mean(avg_fruit) %>% round(2),
            avg_meat = mean(avg_meat) %>% round(2),
            avg_grains = mean(avg_grains) %>% round(2),
            avg_age = mean(AGEC) %>% round(2),
            avg_bmi = mean(BMISC) %>% round(2),
            avg_waist = mean(PHDCMWBC) %>% round(2),
            avg_w2hratio = mean(w2hratio) %>% round(2))

```

Table above shows how many standard deviations away from 0 each cluster is in each variable.

### K-means visualisation

#### Cluster means

```{r}
fit_cluster_kmeans_pca$centers
```

#### Within cluster sum of squares

```{r}
fit_cluster_kmeans_pca$withinss
```
We can see that the clusters have reasonably similar variance

#### Plotting the clusters

```{r}
pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=obese_BMI),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_kmeans,
                        label=cl_kmeans),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("K-means, BMI")


pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=factor(as.character(obese_waist_circum))),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_kmeans,
                        label=cl_kmeans),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("K-means, waist circumference")


pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=obese_w2hratio %>% as.character() %>% as.factor()),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_kmeans,
                        label=cl_kmeans),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none") + ggtitle("K-means, waist to height ratio")
```

```{r}
pcs1 %>% group_by(cl_kmeans) %>% 
  summarise(prop_obese_BMI = sum(as.numeric(obese_BMI))/n())

pcs1 %>% group_by(cl_kmeans) %>% 
  summarise(prop_obese_waist_circum = sum(as.numeric(obese_waist_circum))/n())


pcs1 %>% group_by(cl_kmeans) %>% 
  summarise(prop_obese_w2hratio = sum(as.numeric(obese_w2hratio))/n())
```


```{r}
scaled_numerics <- scale(numerics)
scaled_numerics <- as.data.frame(scaled_numerics)
scaled_numerics$cl_kmeans <- pcs1$cl_kmeans


scaled_numerics %>% group_by(cl_kmeans) %>% 
  summarise(avg_veges = mean(avg_veges) %>% round(2),
            avg_dairy = mean(avg_dairy) %>% round(2),
            avg_fruit = mean(avg_fruit) %>% round(2),
            avg_meat = mean(avg_meat) %>% round(2),
            avg_grains = mean(avg_grains) %>% round(2),
            avg_age = mean(AGEC) %>% round(2),
            avg_bmi = mean(BMISC) %>% round(2),
            avg_waist = mean(PHDCMWBC) %>% round(2),
            avg_w2hratio = mean(w2hratio) %>% round(2))
```

### SVM
#### Models

```{r}
svm_w2hratio <- svm(factor(obese_w2hratio) ~., data = final)
svm_circumf <- svm(factor(obese_waist_circum) ~ ., data=final)
summary(svm_w2hratio)
summary(svm_circumf)
```


```{r}
plot(svm_circumf, data=final, avg_fruit ~ avg_meat)
```



































