---
title: "Plan"
author: "SID: 490300286"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---


# summary of plan

we investigate the effect of diet composition on different measures of obesity.


Obesity can be measured in multiple ways

* BMI
* Waist circumference (https://www.nature.com/articles/s41574-019-0310-7)
* Waist to height ratio (https://bmcmedicine.biomedcentral.com/articles/10.1186/s12916-014-0207-1)

But each of these may categorise obesity differently.

In terms of diet composition, we look at 

* daily servings of meat
* daily servings of veg
* daily servings of dairy
* daily servings of fruit
* daily servings of grains

we then want to see what is the effect of each food type on a measurement for obesity (whether it is significant or not), and compare how this varies across each obesity measurement.

> e.g. Maybe meat servings is a significant contributor to BMI but not to waist circumference.

Then, we can compare to see the effect of each food type across different obesity measures. Examples

> e.g. Does fruit servings have a bigger effect on BMI, waist to height ratio, or waist circumference
> which of the food types has the biggest contribution to BMI reduction/gain.

We can do this by calculating and comparing the standardized betas.

(we can adjust for obvious contributors to each measurement such as age and sex). 


# exec summary of findings

- dairy is insignificant conributor to all measures of obesity


```{r setup, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(factoextra)
library(knitr)
library(kableExtra)
library(gridExtra)
library(broom)
library(fpc)
library(sjPlot)
library(caret)
library(ggfortify)
library(ggforce)
library(Rtsne)
set.seed(2000)
load('tech_data_biom_and_nutr.Rdata')
```


```{r}
# selecting macros
foods = tech_nutr %>% dplyr::select(ABSPID,
                             VEGLEG1N, 
                             VEGLEG2N,
                              FRUIT1N,
                              FRUIT2N,
                              DAIRY1N,
                              DAIRY2N,
                              MEAT1N,
                              MEAT2N,
                              GRAINS1N,
                              GRAINS2N)

# gettin average of macros
avg_veges <- rowMeans(foods[ , c(2,3)], na.rm=TRUE)
avg_fruit <- rowMeans(foods[ , c(4,5)], na.rm=TRUE)
avg_dairy <- rowMeans(foods[ , c(6,7)], na.rm=TRUE)
avg_meat <- rowMeans(foods[ , c(8,9)], na.rm=TRUE)
avg_grains <- rowMeans(foods[ , c(10,11)], na.rm=TRUE)


dat <- cbind(tech_nutr$ABSPID, 
             avg_veges,
             avg_fruit,
             avg_dairy,
             avg_meat,
             avg_grains)

dat <- as_tibble(dat)
colnames(dat)[1] <- "ABSPID"

dat$avg_veges <- as.numeric(dat$avg_veges)
dat$avg_fruit <- as.numeric(dat$avg_fruit)
dat$avg_dairy <- as.numeric(dat$avg_dairy)
dat$avg_meat <- as.numeric(dat$avg_meat)
dat$avg_grains <- as.numeric(dat$avg_grains)



tech_biom1 = tech_biom %>% dplyr::select(c(1:53))

final = merge(dat, tech_biom1, by = "ABSPID")
final = final %>% dplyr::select(avg_veges,
                         avg_fruit,
                         avg_dairy,
                         avg_meat,
                         avg_grains,
                         BMISC,
                         AGEC,
                         SEX,
                         PHDKGWBC,
                         PHDCMHBC,
                         PHDCMWBC,
                         SF2SA1QN)


final$w2hratio = (final$PHDCMWBC)/(final$PHDCMHBC)

# to make results more robust, focus on adults only

final <- final %>% filter(AGEC > 17)

final = final %>% na.omit()
```


## IDA on varibales in df

```{r, message = F, warning = F}
numeric_hist <- function(data, x) {
  ggplot(data, aes_string(x = `x`)) +
  geom_histogram()
}


numeric_hist(final, x = "BMISC")
numeric_hist(final, x = "avg_dairy")
numeric_hist(final, x = "avg_fruit")
numeric_hist(final, x = "avg_meat")
numeric_hist(final, x = "avg_grains")
numeric_hist(final, x = "avg_veges")
numeric_hist(final, x = "AGEC")
```


# short analysis


## linear regression with BMI as outcome

```{r}
bmi_full <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
bmi_null <- lm(BMISC ~ 1, dat = final)

```


## linear regression with waist to heigh ratio as outcome

```{r}
w2hratio_full <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final)
w2hratio_null <- lm(w2hratio ~ 1, dat = final)
```

## linear regression with waist circumference as outcome

```{r}
waist_full <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = final) 
waist_null <- lm(PHDCMWBC ~ 1, dat = final) 

```


NOTE: robust regression was attempted and gave similar results to OLS, so just using OLS instead.


## table of all models

```{r}
tab_model(bmi_full, waist_full, w2hratio_full, show.ci = F, show.se = T)
```


dairy is insignificant for all


## Model selection and performance (10 fold CV)

Model selection using backward and forward selection

### BMI

```{r}
n = 9979

# bmi_AIC is just full model
# bmi_AIC <- step(bmi_full, direction = "backward", trace = F)

bmi_BIC <- step(bmi_full, direction = "backward", trace = F, k = log(n))

# fwd AIC model same as back
# step(bmi_null, scope = list(lower = bmi_null, upper = bmi_full), direction = "forward", trace = F)

```

### waist circumeference

```{r}
waist_AIC <- step(waist_full, direction = "backward", trace = F)

# waist_BIC same as AIC
# waist_BIC <- step(waist_full, direction = "backward", trace = F, k = log(n))


# backward selection same as fwd
# step(waist_null, scope = list(lower = waist_null, upper = waist_full), direction = "forward", trace = F)
```

### w2h ratio


```{r}
w2hratio_AIC <- step(w2hratio_full, direction = "backward", trace = F)
w2hratio_BIC <- step(w2hratio_full, direction = "backward", trace = F, k = log(n))

# backward selection same as fwd
# step(w2hratio_null, scope = list(lower = w2hratio_null, upper = waist_full), direction = "forward", trace = F)
```




```{r}
params = trainControl(method = "cv", number = 10, verboseIter = FALSE)

set.seed(2021)

cv_objects = list(
  bmi_full = train(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat +
    avg_grains + AGEC + SEX, 
    method = "lm", 
    data = final,
    trControl = params),
  bmi_BIC = train(BMISC ~ avg_veges + avg_fruit + avg_meat + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),
  waist_full = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),  
  waist_AIC = train(PHDCMWBC ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params), 
  w2hratio_full = train(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_AIC = train(w2hratio ~ avg_veges + avg_fruit + avg_meat + avg_grains + AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params),   
  w2hratio_BIC = train(w2hratio ~ avg_veges + avg_fruit + avg_grains + 
    AGEC + SEX,
    method = "lm", 
    data = final,
    trControl = params)   
)

cv_results_bmi = resamples(cv_objects[1:2], metric = "Rsquared")
ggplot(cv_results_bmi) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")

cv_results_waist = resamples(cv_objects[3:4], metric = "Rsquared")
ggplot(cv_results_waist) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")


cv_results_w2hratio = resamples(cv_objects[5:7], metric = "Rsquared")
ggplot(cv_results_w2hratio) +
  theme_bw() +
  labs(x = "Models", y = "Mean Absolute Error", title = "10-Fold CV Performance")
```

Regression assumptions

```{r}
model <- cv_objects[2]
#tab_model(model)
```


## comaring standardized betas

```{r}
# creating scaled df
sex <- final$SEX
socioeconomic <- final$SF2SA1QN
scaled_df <- final %>% 
  dplyr::select(-c(SEX, SF2SA1QN)) %>% 
  scale() %>% as.data.frame()


scaled_df$SEX <- sex
scaled_df$SF2SA1QN <- socioeconomic

# lm scaled to get standardized betas to compare
bmi_scaled <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
w2h_scaled <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)
waist_scaled <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC + SEX, dat = scaled_df)


tab_model(bmi_scaled, w2h_scaled, waist_scaled)
```
meat has the biggest effect on BMI overall effect and increases BMI.
fruit and veges decrease waist and waist to height ratio.

# pca and clustering to see any trends in obesity


```{r}

food_groups <- final %>% dplyr::select(avg_dairy,
                                avg_fruit,
                                avg_grains,
                                avg_meat,
                                avg_veges)

numerics <- final %>% dplyr::select(-SEX,
                                    -SF2SA1QN)

# principal component analysis
res.pca <- prcomp(
  food_groups,
  center = T,
  scale = T
)

# extracting pc1 and pc2
pcs = res.pca$x[, c(1,2)] %>% as.data.frame()

```


```{r}
## keeping original data
pcs1 <- pcs
## Creating k-means clustering model
fit_cluster_kmeans_pca <- kmeans(scale(pcs), 5)
# Assigning the result to the data used to create the tsne
pcs1$cl_kmeans <- factor(fit_cluster_kmeans_pca$cluster)
# Creating hierarchical cluster model
fit_cluster_hierarchical_pca <- hclust(dist(scale(pcs)), method = "ward.D2")

plot(fit_cluster_hierarchical_pca, hang = -1)
#  Assigning the result to the data used 
pcs1$cl_hierarchical <- factor(cutree(fit_cluster_hierarchical_pca, k=5))


pcs1$BMISC <- final$BMISC
pcs1$obese <- pcs1$BMISC >= 30


pcs1 %>%
  ggplot(aes(x = PC1,
             y = PC2))+
  geom_point(aes(color=obese),
             alpha = 0.5)+
  scale_color_manual(values = c("#964B00", "#8F00FF",  "#0000FF", "#00FF00", 
                                "#FFA500", "#808080", "#FF0000")) +
    geom_mark_ellipse(aes(color = cl_hierarchical,
                        label=cl_hierarchical),
                    expand = unit(0.5,"mm"),
                    label.buffer = unit(-5, 'mm')) +
  theme(legend.position = "none")



pcs1 %>% group_by(cl_hierarchical) %>% 
  summarise(prop_obese = sum(as.numeric(obese))/n())
```


```{r}
scaled_numerics <- scale(numerics)
scaled_numerics <- as.data.frame(scaled_numerics)
scaled_numerics$cl_hierarchical <- pcs1$cl_hierarchical


scaled_numerics %>% group_by(cl_hierarchical) %>% 
  summarise(avg_veges = mean(avg_veges) %>% round(2),
            avg_dairy = mean(avg_dairy) %>% round(2),
            avg_fruit = mean(avg_fruit) %>% round(2),
            avg_meat = mean(avg_meat) %>% round(2),
            avg_grains = mean(avg_grains) %>% round(2),
            avg_age = mean(AGEC) %>% round(2),
            avg_bmi = mean(BMISC) %>% round(2),
            avg_waist = mean(PHDCMWBC) %>% round(2),
            avg_w2hratio = mean(w2hratio) %>% round(2))

```

Table above shows how many standard deviations away from 0 each cluster is in each variable.






