---
title: "Plan"
author: "SID: 490300286"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---


# summary of plan

we investigate the effect of diet composition on different measures of obesity.


Obesity can be measured in multiple ways

* BMI
* Waist circumference (https://www.nature.com/articles/s41574-019-0310-7)
* Waist to height ratio (https://bmcmedicine.biomedcentral.com/articles/10.1186/s12916-014-0207-1)

But each of these may categorise obesity differently.

In terms of diet composition, we look at 

* daily servings of meat
* daily servings of veg
* daily servings of dairy
* daily servings of fruit
* daily servings of grains

we then want to see what is the effect of each food type on a measurement for obesity (whether it is significant or not), and compare how this varies across each obesity measurement.

> e.g. Maybe meat servings is a significant contributor to BMI but not to waist circumference.

Then, we can compare to see the effect of each food type across different obesity measures. Examples

> e.g. Does fruit servings have a bigger effect on BMI, waist to height ratio, or waist circumference
> which of the food types has the biggest contribution to BMI reduction/gain.

We can do this by calculating and comparing the standardized betas.

(we can adjust for obvious contributors to each measurement such as age and sex). 


# short analysis

```{r setup, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(factoextra)
library(knitr)
library(kableExtra)
library(gridExtra)
library(broom)
library(fpc)
library(sjPlot)
library(QuantPsyc)
set.seed(2000)
load('tech_data_biom_and_nutr.Rdata')
```


```{r}
# selecting macros
foods = tech_nutr %>% dplyr::select(ABSPID,
                             VEGLEG1N, 
                             VEGLEG2N,
                              FRUIT1N,
                              FRUIT2N,
                              DAIRY1N,
                              DAIRY2N,
                              MEAT1N,
                              MEAT2N,
                              GRAINS1N,
                              GRAINS2N)

# gettin average of macros
avg_veges <- rowMeans(foods[ , c(2,3)], na.rm=TRUE)
avg_fruit <- rowMeans(foods[ , c(4,5)], na.rm=TRUE)
avg_dairy <- rowMeans(foods[ , c(6,7)], na.rm=TRUE)
avg_meat <- rowMeans(foods[ , c(8,9)], na.rm=TRUE)
avg_grains <- rowMeans(foods[ , c(10,11)], na.rm=TRUE)


dat <- cbind(tech_nutr$ABSPID, 
             avg_veges,
             avg_fruit,
             avg_dairy,
             avg_meat,
             avg_grains)

dat <- as_tibble(dat)
colnames(dat)[1] <- "ABSPID"

dat$avg_veges <- as.numeric(dat$avg_veges)
dat$avg_fruit <- as.numeric(dat$avg_fruit)
dat$avg_dairy <- as.numeric(dat$avg_dairy)
dat$avg_meat <- as.numeric(dat$avg_meat)
dat$avg_grains <- as.numeric(dat$avg_grains)



tech_biom1 = tech_biom %>% dplyr::select(c(1:53))

final = merge(dat, tech_biom1, by = "ABSPID")
final = final %>% dplyr::select(avg_veges,
                         avg_fruit,
                         avg_dairy,
                         avg_meat,
                         avg_grains,
                         BMISC,
                         AGEC,
                         SEX,
                         PHDKGWBC,
                         PHDCMHBC,
                         PHDCMWBC,
                         SF2SA1QN)


final$w2hratio = (final$PHDCMWBC)/(final$PHDCMHBC)

final = final %>% na.omit()
```


## linear regression with BMI as outcome


```{r}
bmi <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final)
tab_model(bmi)
```


## linear regression with waist to heigh ratio as outcome

```{r}
w2hratio <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final)
tab_model(w2hratio)
```


## linear regression with waist circumference as outcome

```{r}
waist <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final) 
tab_model(waist)
```


e.g. looking at above formulas, we can see 

* when BMI is the outcome, dairy is not significant, veges is not significant (almost is though)
* when waist-to-height ratio is the ooutcome, dairy and meat are not significant
* when waist circumference is the outcome, veges and grains are not significant.


## comaring standardized betas

```{r}
bmi_betas <- QuantPsyc::lm.beta(bmi)
w2h_betas <- QuantPsyc::lm.beta(w2hratio)
waist_betas <- QuantPsyc::lm.beta(waist)

rbind(bmi_betas,
      w2h_betas,
      waist_betas) %>% round(2) %>% kable()

```
meat has the biggest effect on BMI overall effect and increases BMI.
fruit and veges decrease waist and waist to height ratio.



