---
title: "Model Assumptions"
author: "SID: 490419348"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
---

#### Libraries
```{r, warning=FALSE, message=FALSE}
# Load in all libraries and seeds
library(tidyverse)
library(here)      # directory referencing
library(readxl)    # reading Excel files
library(janitor)   # data cleaning 
library(stringr)   # string manipuation
library(tidyr)     # new tidy functions
library(caret)
library(grid)
library(dplyr)
library(gridExtra)
library(mlbench)
library(naniar)
library(broom)
library(knitr)
library(sjPlot)
set.seed(69)
```




### Loading and cleaning data
#### Loading tech data
```{r}
load('../data/tech_data_biom_and_nutr.Rdata')
```

```{r}
# selecting macros
foods = tech_nutr %>% dplyr::select(ABSPID,
                             VEGLEG1N, 
                             VEGLEG2N,
                              FRUIT1N,
                              FRUIT2N,
                              DAIRY1N,
                              DAIRY2N,
                              MEAT1N,
                              MEAT2N,
                              GRAINS1N,
                              GRAINS2N)

# gettin average of macros
avg_veges <- rowMeans(foods[ , c(2,3)], na.rm=TRUE)
avg_fruit <- rowMeans(foods[ , c(4,5)], na.rm=TRUE)
avg_dairy <- rowMeans(foods[ , c(6,7)], na.rm=TRUE)
avg_meat <- rowMeans(foods[ , c(8,9)], na.rm=TRUE)
avg_grains <- rowMeans(foods[ , c(10,11)], na.rm=TRUE)


dat <- cbind(tech_nutr$ABSPID, 
             avg_veges,
             avg_fruit,
             avg_dairy,
             avg_meat,
             avg_grains)

dat <- as_tibble(dat)
colnames(dat)[1] <- "ABSPID"

dat$avg_veges <- as.numeric(dat$avg_veges)
dat$avg_fruit <- as.numeric(dat$avg_fruit)
dat$avg_dairy <- as.numeric(dat$avg_dairy)
dat$avg_meat <- as.numeric(dat$avg_meat)
dat$avg_grains <- as.numeric(dat$avg_grains)



tech_biom1 = tech_biom %>% dplyr::select(c(1:53))

final = merge(dat, tech_biom1, by = "ABSPID")
final = final %>% dplyr::select(avg_veges,
                         avg_fruit,
                         avg_dairy,
                         avg_meat,
                         avg_grains,
                         BMISC,
                         AGEC,
                         SEX,
                         PHDKGWBC,
                         PHDCMHBC,
                         PHDCMWBC,
                         SF2SA1QN)


final$w2hratio = (final$PHDCMWBC)/(final$PHDCMHBC)

final = final %>% na.omit()
```

## linear regression with BMI as outcome


```{r}
bmi <- lm(BMISC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final)
tab_model(bmi)
```


## linear regression with waist to heigh ratio as outcome

```{r}
w2hratio <- lm(w2hratio ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final)
tab_model(w2hratio)
```


## linear regression with waist circumference as outcome

```{r}
waist <- lm(PHDCMWBC ~ avg_veges + avg_fruit + avg_dairy + avg_meat + avg_grains + AGEC, dat = final) 
tab_model(waist)
```


e.g. looking at above formulas, we can see 

* when BMI is the outcome, dairy is not significant, veges is not significant (almost is though)
* when waist-to-height ratio is the ooutcome, dairy and meat are not significant
* when waist circumference is the outcome, veges and grains are not significant.








